\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing economic inequality with tax data - Switzerland from 1945 to 2010}

### 2014-06-30 paper/subdocuments/results.Rnw
noch mehr aufgeräumt

<<setup_gini_data>>=
library(foreign)
library(ggplot2)
df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
df$zeros <- df$null_norm/df$cpop # i only use normal cases here
@

@@

<<reldist20032010>>=
library(reldist, quietly=TRUE, warn.conflicts = FALSE)

@@

<<polindex20032010>>=
wgt<-rep(1,length(bd.2003.ind))
format(rpy(y=bd.2010.ind,yo=bd.2003.ind,
           ywgt=wgt,yowgt=wgt,pvalue=FALSE),
       digits=3) # Median Index, middle value is the estimate 
format(rpluy(y=bd.2010.ind,yo=bd.2003.ind,
             ywgt=wgt,yowgt=wgt,pvalue=FALSE),
       digits=3) # Lower Index
format(rpluy(y=bd.2010.ind,yo=bd.2003.ind,
             ywgt=wgt,yowgt=wgt,pvalue=FALSE,
             upper=TRUE),
       digits=3) # Upper index

gini(bd.2010.ind)-gini(bd.2003.ind)

@@

% Variante 1993/1994
<<setup_data_ESTVspecialcases9394, echo=FALSE,warning=FALSE,hide=TRUE>>=
library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
library(ggplot2)
#library(reldist)
normal <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
normal <- select(normal, -G_reink,-G_taxed,-ppop0,-G_steink0)
#normal <- normal %.% select(-G_reink, -G_taxed, -ppop0, -G_steink0)
pooled <- read.csv("../../data/ginis_und_perzentile_beides.csv",sep="\t")
#special <- read.csv("data/ginis_und_perzentile_sonder.csv",segep="\t")
normal$case <- "normal case"
#special$case <- "special case"
pooled$case <- "pooled, normal and special cases"
d<-rbind(normal,pooled)
d<-d %.% 
  filter(kanton=="CH",steuerperiode==1993.5)

@@

<<specialcases9394, echo=FALSE,warning=FALSE,hide=TRUE>>=
#plot relative distribution

@@

<<polindexspecialcases9394>>=
wgt<-rep(1,length(normal_i))
format(rpy(y=pooled_i,yo=normal_i,
           ywgt=wgt,yowgt=wgt,pvalue=FALSE),
       digits=3) # Median Index, middle value is the estimate 
format(rpluy(y=pooled_i,yo=normal_i,
             ywgt=wgt,yowgt=wgt,pvalue=FALSE),
       digits=3) # Lower Index
format(rpluy(y=pooled_i,yo=normal_i,
             ywgt=wgt,yowgt=wgt,pvalue=FALSE,
             upper=TRUE),
       digits=3) # Upper index

gini(pooled_i)-gini(normal_i)
@

@@

% Variante 2010
<<setup_data_specialcases10, echo=FALSE,warning=FALSE,hide=TRUE>>=
library(foreign)
library(ggplot2)
#library(reldist)
normal <- filter(read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t"),kanton=="CH",steuerperiode==2010)
start <- which(names(normal)=="p1")
end <- which(names(normal)=="p99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,99)
normal <- data.frame(t(normal[1,start:end]),percentile)
names(normal)[1]=c("inc")

@@

<<specialcases10, echo=FALSE,warning=FALSE,hide=TRUE>>=

@@

<<polindexspecialcases10, echo=FALSE,warning=FALSE,hide=TRUE>>=

wgt<-rep(1,length(normal_i))

format(rpy(y=pooled_i,yo=normal_i,
           ywgt=wgt,yowgt=wgt,pvalue=FALSE),
       digits=3) # Median Index, middle value is the estimate 
format(rpluy(y=pooled_i,yo=normal_i,
             ywgt=wgt,yowgt=wgt,pvalue=FALSE),
       digits=3) # Lower Index
format(rpluy(y=pooled_i,yo=normal_i,
             ywgt=wgt,yowgt=wgt,pvalue=FALSE,
             upper=TRUE),
       digits=3) # Upper index

gini(normal_i)-gini(pooled_i)
@

@@

<<with_without_zeros_RelDist_mitBrülhart,warning=FALSE>>=
bd.2010 <- filter(read.csv("../../data/data_Schweiz.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
bd.0.2010 <- filter(read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
start <- which(names(bd.2010)=="p1")
end <- which(names(bd.2010)=="p99_99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
names(bd.2010)[1]=c("inc")
bd.0.2010<- data.frame(t(bd.0.2010[1,start:end]),percentile)
names(bd.0.2010)[1]=c("inc")

@@

gA0 <- reldist(y=bd.2010.ind, yo=bd.0.2010.ind,
               smooth=0.4, ci=FALSE,
               show="residual",
               entropy=NA,
               bar=TRUE, quiet=FALSE,
               ylim=c(0.5,2), ylab="",
               yolabs=seq(-1,3,by=0.5),
               discrete=FALSE,
               xlab="Proportion of 2010 Distribution (including non-taxed)")
title(main=paste("Effect of different shape"))
abline(h=1,lty=2)

# Polarization index

wgt<-rep(1,length(bd.2010.ind))

format(rpy(y=bd.2010.ind,yo=bd.0.2010.ind,
           ywgt=wgt,yowgt=wgt,pvalue=FALSE),
       digits=3) # Median Index, middle value is the estimate 
format(rpluy(y=bd.2010.ind,yo=bd.0.2010.ind,
             ywgt=wgt,yowgt=wgt,pvalue=FALSE),
       digits=3) # Lower Index
format(rpluy(y=bd.2010.ind,yo=bd.0.2010.ind,
             ywgt=wgt,yowgt=wgt,pvalue=FALSE,
             upper=TRUE),
       digits=3) # Upper index

@@

gA0 <- reldist(y=bd.2010.ind, yo=bd.0.2010.ind,
               smooth=0.4, ci=FALSE,
               show="residual",
               entropy=NA,
               bar=TRUE, quiet=FALSE,
               ylim=c(0.5,2), ylab="",
               yolabs=seq(-1,3,by=0.5),
               discrete=FALSE,
               xlab="Proportion of 2010 Distribution (including non-taxed)")
title(main=paste("Effect of different shape"))
abline(h=1,lty=2)

# Polarization index

wgt<-rep(1,length(bd.2010.ind))

format(rpy(y=bd.2010.ind,yo=bd.0.2010.ind,
           ywgt=wgt,yowgt=wgt,pvalue=FALSE),
       digits=3) # Median Index, middle value is the estimate 
format(rpluy(y=bd.2010.ind,yo=bd.0.2010.ind,
             ywgt=wgt,yowgt=wgt,pvalue=FALSE),
       digits=3) # Lower Index
format(rpluy(y=bd.2010.ind,yo=bd.0.2010.ind,
             ywgt=wgt,yowgt=wgt,pvalue=FALSE,
             upper=TRUE),
       digits=3) # Upper index

@@

#####################################################
library(reshape)
library(dplyr)
try(habe0911<-read.table("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
try(habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
try(habe0911<-read.table("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
try(g2009 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2009_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2010 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2010_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2011 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2011_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2009 <- read.csv("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE2009_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2010 <- read.csv("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE2010_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2011 <- read.csv("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE2011_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
names(g2009)[3:4]<-c("SRH","Gewicht")
names(g2010)[3:4]<-c("SRH","Gewicht")
names(g2011)[3:4]<-c("SRH","Gewicht")
jg091011 <- rbind(g2009,g2010,g2011)
all.equal(match(jg091011$HaushaltID,habe0911$HaushaltID),1:nrow(habe0911))
habe0911$jahresgewicht <- jg091011$Gewicht

@@

<<BrülhartvsHABE,warning=FALSE>>=
library(ggplot2)
library(reldist, quietly=TRUE, warn.conflicts = FALSE)

@@

##
# Unmarried ESTV

estv2010.unmarried<-estv2010[c(18:38),c(1,3)]
names(estv2010.unmarried)[2]=c("unmarried")
##
# Simulate individual data points
# Percentile-Plot (descriptive  purpose)
# cumulative probabilites
cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
# probabilities
prob<-c(cum.p[1],diff(cum.p), 1-0.9999)
# extreme values beyond x (to sample)
#freq<-max(normal$anz_pflichtige)
freq<-3000
init<-0
fin<-(abs(max(estv2010.unmarried$unmarried,na.rm=TRUE))+5)
# generate the sequence to take pairs from
ival<-c(init,estv2010.unmarried$unmarried,fin)
len<-10000 # sequence of each pair
s<-sapply(2:length(ival),function(i){
  seq(ival[i-1],ival[i],length.out=len)
})
s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
estv2010.unmarried.ind<-data.frame(s)
names(estv2010.unmarried.ind)[1]=c("inc")
estv2010.unmarried.ind$inc<-estv2010.unmarried.ind$inc/1000

# Select unmarried - HABE
try(habe0911<-read.table("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
try(habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
habe10<-habe0911 %.% 
  filter(Jahr08==2010)
habe10.unmarried<-habe0911 %.% 
  filter(Familientyp08<140,Jahr08==2010)
habe10.unmarried<-data.frame(habe10.unmarried$VerfuegbaresEinkommen08)
names(habe10.unmarried)<-"inc_habe"
habe10.unmarried$inc_habe[habe10.unmarried$inc_habe<0]<-0 # One values is -20'000
habe10.unmarried$inc_habe<-habe10.unmarried$inc_habe*12/1000 # transform from monthly income to income per year

@@

>>>>

<<setup_gini_data>>=
library(foreign,quietly=TRUE, warn.conflicts=FALSE)
library(ggplot2)
df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
df$zeros <- df$null_norm/df$cpop # i only use normal cases here
@

@@

<<reldist20032010, echo=FALSE>>=
library(reldist, quietly=TRUE, warn.conflicts = FALSE)

@@

<<polindex20032010, results='asis'>>=
# Setup functin to produce latex tables with polarization indices and difference in ginis
polTable <- function(y, yo, ywgt=1, yowgt=1,ndig=2,title="Inequality Indices") {
stopifnot(is.numeric(y), is.numeric(yo)) # all TRUE
#y vector with new distribution
#yo vector with reference distribution
#ywgt yowgt: weights to use (optional)
#ndig: digits to display, default 2
#title: table title (optional)

# load neccessary libraries
library(reldist,quietly=TRUE,warn.conflicts=FALSE) #for Latex tables
library(Hmisc,quietly=TRUE,warn.conflicts=FALSE)

# Create vector of weights 1 if nothing else is defined
if(length(ywgt)==1 & ywgt==1) ywgt <- rep(1, length(y))
if(length(yowgt)==1 & yowgt==1) yowgt <- rep(1, length(yo))

med <- rpy(y=y,yo=yo, ywgt=ywgt,yowgt=yowgt,pvalue=FALSE)[2] # Median Index, middle value is the estimate so we take [2]
lower <- rpluy(y=y,yo=yo, ywgt=ywgt,yowgt=yowgt,pvalue=FALSE)[2] # Lower Index
upper <- rpluy(y=y,yo=yo, ywgt=ywgt,yowgt=yowgt,pvalue=FALSE, upper=TRUE)[2] # Upper index
dgini <- gini(y)-gini(yo)

latex(format(c("Median Index"=med, "Lower Index"=lower, "Upper Index"=upper, "$\\Delta$ Gini"=dgini),digits=ndig),file="",title=title)
}

polTable(bd.2010.ind,bd.2003.ind)
@

@@

% Variante 1993/1994
<<setup_data_ESTVspecialcases9394>>=
library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
library(ggplot2, quietly=TRUE, warn.conflicts = FALSE)
#library(reldist)
normal <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
normal <- select(normal, p1:p99, kanton, steuerperiode, -p96,-p97,-p98)
#normal <- normal %.% select(-G_reink, -G_taxed, -ppop0, -G_steink0)
pooled <- read.csv("../../data/ginis_und_perzentile_beides.csv",sep="\t")
pooled <- select(pooled, kanton, steuerperiode, p1:p99)
#special <- read.csv("data/ginis_und_perzentile_sonder.csv",segep="\t")
normal$case <- "normal case"
#special$case <- "special case"
pooled$case <- "pooled, normal and special cases"
d<-rbind(normal,pooled)
d<-d %.% 
  filter(kanton=="CH",steuerperiode==1993.5)

@@

<<specialcases9394>>=
#plot relative distribution

@@

<<polindexspecialcases9394, results='asis'>>=
polTable(pooled_i,normal_i)
@

@@

% Variante 2010
<<setup_data_specialcases10>>=
library(foreign)
library(ggplot2)
#library(reldist)
normal <- filter(read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t"),kanton=="CH",steuerperiode==2010)
normal <- select(normal, p1:p99, kanton, steuerperiode, -p96,-p97,-p98)
start <- which(names(normal)=="p1")
end <- which(names(normal)=="p99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,99)
normal <- data.frame(t(normal[1,start:end]),percentile)
names(normal)[1]=c("inc")

@@

<<specialcases10>>=

@@

<<polindexspecialcases10, results='asis'>>=
polTable(pooled_i,normal_i)
@

@@

<<with_without_zeros_RelDist_mitBruelhart,warning=FALSE>>=
bd.2010 <- filter(read.csv("../../data/data_Schweiz.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
bd.0.2010 <- filter(read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
start <- which(names(bd.2010)=="p1")
end <- which(names(bd.2010)=="p99_99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
names(bd.2010)[1]=c("inc")
bd.0.2010<- data.frame(t(bd.0.2010[1,start:end]),percentile)
names(bd.0.2010)[1]=c("inc")

@@

gA0 <- reldist(y=bd.2010.ind, yo=bd.0.2010.ind,
               smooth=0.4, ci=FALSE,
               show="residual",
               entropy=NA,
               bar=TRUE, quiet=FALSE,
               ylim=c(0.5,2), ylab="",
               yolabs=seq(-1,3,by=0.5),
               discrete=FALSE,
               xlab="Proportion of 2010 Distribution (including non-taxed)")
title(main=paste("Effect of different shape"))
abline(h=1,lty=2)
@

<<polindexzeros, results='asis'>>=
polTable(bd.2010.ind,bd.0.2010.ind)
@

@@

gA0 <- reldist(y=bd.2010.ind, yo=bd.0.2010.ind,
               smooth=0.4, ci=FALSE,
               show="residual",
               entropy=NA,
               bar=TRUE, quiet=FALSE,
               ylim=c(0.5,2), ylab="",
               yolabs=seq(-1,3,by=0.5),
               discrete=FALSE,
               xlab="Proportion of 2010 Distribution (including non-taxed)")
title(main=paste("Effect of different shape"))
abline(h=1,lty=2)
@

<<polindexzeros2, results='asis'>>=
polTable(bd.2010.ind,bd.0.2010.ind)
@

@@

#####################################################
library(reshape, quietly=TRUE, warn.conflicts=FALSE)
library(dplyr, quietly=TRUE, warn.conflicts=FALSE)
try(habe0911<-read.table("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
try(habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
try(habe0911<-read.table("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
try(g2009 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2009_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2010 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2010_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2011 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2011_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2009 <- read.csv("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE2009_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2010 <- read.csv("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE2010_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2011 <- read.csv("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE2011_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2009 <- read.csv("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE2009_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2010 <- read.csv("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE2010_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
try(g2011 <- read.csv("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE2011_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
names(g2009)[3:4]<-c("SRH","Gewicht")
names(g2010)[3:4]<-c("SRH","Gewicht")
names(g2011)[3:4]<-c("SRH","Gewicht")
jg091011 <- rbind(g2009,g2010,g2011)
#all.equal(match(jg091011$HaushaltID,habe0911$HaushaltID),1:nrow(habe0911)) ## proves it is ok to just attach weights in the default order
habe0911$jahresgewicht <- jg091011$Gewicht

@@

<<BruelhartvsHABE,warning=FALSE>>=
library(ggplot2)
library(reldist, quietly=TRUE, warn.conflicts = FALSE)

@@

##
# Unmarried ESTV
estv2010 <- read.csv("../../data/estv_normal_sonder_mitNuller_stE_2010.csv",sep=";")
estv2010.unmarried<-estv2010[c(18:38),c(1,3)]
names(estv2010.unmarried)[2]=c("unmarried")
##
# Simulate individual data points
# Percentile-Plot (descriptive  purpose)
# cumulative probabilites
cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
# probabilities
prob<-c(cum.p[1],diff(cum.p), 1-0.9999)
# extreme values beyond x (to sample)
#freq<-max(normal$anz_pflichtige)
freq<-3000
init<-0
fin<-(abs(max(estv2010.unmarried$unmarried,na.rm=TRUE))+5)
# generate the sequence to take pairs from
ival<-c(init,estv2010.unmarried$unmarried,fin)
len<-10000 # sequence of each pair
s<-sapply(2:length(ival),function(i){
  seq(ival[i-1],ival[i],length.out=len)
})
s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
estv2010.unmarried.ind<-data.frame(s)
names(estv2010.unmarried.ind)[1]=c("inc")
estv2010.unmarried.ind$inc<-estv2010.unmarried.ind$inc/1000

# Select unmarried - HABE
habe10<-habe0911 %.% 
  filter(Jahr08==2010)
habe10.unmarried<-habe0911 %.% 
  filter(Familientyp08<140,Jahr08==2010)
habe10.unmarried<-data.frame(habe10.unmarried$VerfuegbaresEinkommen08)
names(habe10.unmarried)<-"inc_habe"
habe10.unmarried$inc_habe[habe10.unmarried$inc_habe<0]<-0 # One values is -20'000
habe10.unmarried$inc_habe<-habe10.unmarried$inc_habe*12/1000 # transform from monthly income to income per year

@@

#1918 fixen
estv <- within(estv, {
    steink[steuerperiode==1918]<-(eink_ll[steuerperiode==1918]+eink_ul[steuerperiode==1918])/2*anz_pflichtige[steuerperiode==1918]
})
